### 2015-02-10 paper/experiments.tex
rewritting those chapters

We conducted several performance tests to measure the throughput and scalability of our system. All these tests were performed on a LAN system called Lattice which has a network bandwidth of 1 Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 16 GB of memory. We used Apache S4 0.6.0 version and Apache storm 0.9.2 version. Sample code used for all tests can be found here\cite{solutionCode}.

\subsection{Inter node communication}
We used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. Our solution \textit{EventProducer} and Yahoo S4\cite{neumeyer2010s4} \textit{EventProducer} read a file containing over 7500000 ECG records and push events with multiple threads. Twitter Storm\cite{twitterStorm} does not allow user threads to push data. So we used the \textit{spout} thread to send messages into the system. \textit{EventReceiver}  receives the ECG events, processes them and calculates heart rate interval periodically.

\subsubsection{Throughput}
We executed our system with 1, 2 and 4  \textit{EventReceivers} for each system and measured the throughput, load average and network bandwidth for each case. Throughput was measured at the \textit{EventProducer} by calculating the total time required to send messages and the total number of messages send. Load average and network bandwidth were measured using top and atop linux commands respectively. Figure \ref{throuput} shows the throughput variation with the number of nodes for all systems.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{ecgGraph.png}
        \caption{ECG Process Graph}
        \label{ecgGraph}
\end{figure}
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{throughput.png}
        \caption{Throughput of the systems}
        \label{throuput}
\end{figure}

As shown in the Figure \ref{throuput}, our solution outperforms Twitter Storm \cite{twitterStorm} and Yahoo S4 \cite{neumeyer2010s4}. However by looking at the Figure \ref{throuput}, one might think even our solution is not scalable although it performs well. In order to examine the reason behind this, we looked into the network bandwidth and CPU load average values.

Figure \ref{networkandload} shows the network bandwidth usage and the CPU load average at each node. For multiple node scenarios, since we observed the same values for both bandwidth usage and CPU load average at each receiver, we took the average values of them.

\begin{figure*}[!t]
	\centering
	\subfloat[Network Usage]{\includegraphics[width=3.0in]{network.png}}
	\hfil
	\subfloat[CPU Load Average]{\includegraphics[width=3.0in]{loadaverage.png}}
	\caption{Network usage and CPU Load Average of the System}
	\label{networkandload}
\end{figure*}

By looking at the graphs, first it can be observed that our solution utilities all available network bandwidth (98\%) even with two receivers. In the one receiver case, high CPU load average at the receiver indicates that it has utilized available CPU power. Adding two receivers has increased the CPU power at receiving side utilizing all available network bandwidth. Therefore, we can not increase the throughput without increasing network bandwidth although more CPU capacity is available at the receiver nodes. Having more receiver nodes, we can observe that network bandwidth and CPU load average is reduced at each receiver, since load is shared among the receivers. However for other two systems, it neither hits the maximum network bandwidth available nor the full CPU power available at any stage. If a system does not make efficient use of the available resources then adding more resources will not scale up the system. As explained in earlier, our solution utilities all available resources by using parallel TCP connections and having efficient message transformations.

\subsubsection{Efficiency of Message Serialization}
After measuring the throughput, we examined the  total amount of extra bytes each system sends to transfer the information from \textit{EventProducer} to \textit{EventReceiver}. Our ECGEvent has two double fields called time and value (ecg signal value) and a streamID which is a 4 byte string. So we calculated total bytes to send this information as 20 bytes (16 for two doubles and 4 for streamID). Then by using the throughput and the bandwidth usage of the system, we calculated the total number of bytes each system uses at network layer to send this message. Finally we calculated the overhead bytes for each system by reducing 20 information bytes. Figure \ref{efficiency} compares these numbers.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{efficiency.png}
        \caption{Efficiency of Message Serialization}
        \label{efficiency}
\end{figure}

As shown in the figure \ref{efficiency}, Twitter Storm \cite{twitterStorm} uses a minimum over head while Yahoo S4 \cite{neumeyer2010s4} performs worse compared to other two. This overhead is due to its usage of Java serialization. We analyzed our solutions' extra overhead of 32 bytes. Our solution adds a sequence number (an integer), receiving process id (a 8 byte string for this scenario), sending process id (a 8 byte string for this scenario) to order messages, dispatch the message to correct process, and to parse the message at the server. Altogether it adds 26 (serialization format of a string uses two more bytes to keep the string length) extra bytes including 2 overhead bytes for streamID, to each message at application level and another 4 bytes due to TCP overhead. This is the trade off we had to pay to improve the parallelism compared to direct one TCP communication per process.

\subsubsection{Latency}
We examined the message latency between the event producer and receiver. For this experiment, we sent the timestamp at the producer along the message and calculated the latency at the receiver. We took 6000 samples over the experiment duration. As shown in the figure \ref{latancydis}, we could observe the mean latency of 185.05ms with the 95\% confidence interval of (182.56ms, 187.54ms).

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{latencyHistogram.png}
        \caption{Message Latency Distribution}
        \label{latancydis}
\end{figure}

\subsection{Scalability}
We performed a scalability test for our system using the 3-level graph shown in Figure \ref{multigraph}. The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}. This project explores the possibility of using stream processing to process data generated in the smart grid domain. The data is generated by sensors attached to a device called “smart plug” which is plugged as an intermediate layer between a power outlet and the plug of a home appliance. The sensors capture various measures related to power consumption of each connected device and send these measurements continuously to a central processing system. Each household contains several of such smart plugs which will report the sensor readings once in every second. At the central processing system, these data should be processed in order to forecast the short time load and to generate load statistics for real time demand management. We implemented the first part of this project which predicts the load using a machine learning algorithm. This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices. We used the publicly available 500MB of data to generate the events.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{multigraph.png}
        \caption{Multilevel Node Graph}
        \label{multigraph}
\end{figure}

@@

As in the earlier case, we conducted our experiments using one producer and incrementing the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer. We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively. We measured the throughput,  network bandwidth and the CPU load average at the producer to examine the scalability of the system. Since the one receiver unit throughput is greater than that of the Twitter Storm one node scenario, we only used our system for this experiment.  Results are shown in the Figure \ref{scalability}.

@@

For one the receiver unit case, we observed very high load average and network bandwidth usage at \textit{avgCal} node since it sends messages to two nodes. As shown in  Figure \ref{scalability}, we could linearly scale up the system by adding more receiver units to add more cpu power to the system. When increasing the receiver units, we could observe throughput of the system increases proportional to number of receiving units. Our CPU load average of 13 at the producer indicates it has utilized all available CPU power and it is required to add more producers to scale this system further up.

>>>>

We conducted several performance benchmarks to measure the throughput and scalability of our system. All these tests were performed on a LAN system called Lattice which has a network bandwidth of 1 Gbps. All nodes are Intel(R) Xeon(R) 2.4GHz 4 core duo machines with 16 GB of memory. We used Apache S4 0.6.0 version and Apache storm 0.9.2 version. Sample code used for all tests can be found here \cite{solutionCode}.

\subsection{Effect of message buffer size and thread pool}
For this experiment, we used a graph as shown in Figure \ref{ecgGraph} to process ECG signal data. The EventProducer reads a file containing over 7500000
 ECG records and pushes events with multiple threads. EventReceiver receives the ECG events, processes them and calculates heart rate interval periodically using Pan Tompkins algorithm.

\begin{figure}[!t]
        \centering
        \includegraphics[width=2.0in]{ecgGraph.png}
        \caption{ECG Process Graph}
        \label{ecgGraph}
\end{figure}

We measured the throughput between two nodes for different thread pool and message buffer sizes to understand the effect of those parameters to throughput. As shown in figure \ref{throughputvar} throughput increases clearly with the message buffer size and increase with thread pool size for larger message buffer sizes. Therefore we can conclude both make a positive impact on the throughput.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{throughputvar.png}
        \caption{Throughput Variation}
        \label{throughputvar}
\end{figure}

\subsection{Load balancing}

We used the same graph (Figure \ref{ecgGraph}) used in above experiment with 1,2 and 4 \textit{EventReciver} nodes with one \textit{EventSender} node for this experiment. Since 40 thread pool size and 40 message buffer size gives highest throughput, we used that values with our solution. Further we compared our system with Yahoo S4 and Twitter Storm with the same graph. Twitter Storm places tasks arbitrarily within the available nodes. Therefore we used 1 task for \textit{EventSender} and 1,2 and 4 tasks for \textit{EventReceivers} to place them in different nodes. This makes Twitter Storm execute only one thread to process data in each node. In order to compare this situation, we measured our system performance with 1 thread and 10 message buffer size as well.

Figure \ref{throuput} shows the throughput of the systems. All systems keep initial throughput when increasing nodes. But our solution out performs all the other systems. Figure \ref{networkusage} shows the network bandwidth at each node (measured using atop linux command). For each framework sending node consumes same amount of network bandwidth since it sends all messages to other nodes. For receiving side each receiving node consumes corresponding portion of the bandwidth according to number of receiving nodes. For an example for 4 node configuration each receiving node consumes only one fourth of the sending node bandwidth. Compared to other systems our solution consumes all the available bandwidth.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{throughput.png}
        \caption{Throughput of the systems}
        \label{throuput}
\end{figure}

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{network.png}
        \caption{Network Usage of the Systems}
        \label{networkusage}
\end{figure}

\subsection{Multistage Graph Processing}
We performed a multistage graph test for our system using the 3-level graph shown in Figure \ref{multigraph}. The data as well as the processing logic were obtained from the Grand Challenge problem at 8th ACM International Conference on Distributed Event Based Systems \cite{acm}. This project explores the possibility of using stream processing to process data generated in the smart grid domain. The data is generated by sensors attached to a device called “smart plug” which is plugged as an intermediate layer between a power outlet and the plug of a home appliance. The sensors capture various measures related to power consumption of each connected device and send these measurements continuously to a central processing system. Each household contains several of such smart plugs which will report the sensor readings once in every second. At the central processing system, these data should be processed in order to forecast the short time load and to generate load statistics for real time demand management. We implemented the first part of this project which predicts the load using a machine learning algorithm. This algorithm divides the events to fixed intervals called  time slices and uses average value of time slices to predict the average of feature time slices. We used the publicly available 500MB of data to generate the events.

\begin{figure}[!t]
        \centering
        \includegraphics[width=2.0in]{multigraph.png}
        \caption{Multilevel Node Graph}
        \label{multigraph}
\end{figure}

@@

As in the earlier case, we conducted our experiments using one producer and incrementing the other processing nodes by 1, 2 and 4 times to measure the throughput increase at the producer. We ran each node in a separate machine so that our receiver configurations used 5, 10, 15 machines respectively. All nodes were configured to use both thread pool and message buffer size of 40. We measured the throughput, network bandwidth and the CPU load average at the producer to examine the scalability of the system. Since the one receiver unit throughput is greater than that of the Twitter Storm one node scenario, we only used our system for this experiment. As shown in Figure \ref{scalability} system scales up until it fully utilized the available network bandwidth.

@@

\subsection{Scalability and Latency}
We analyze the scalability of our framework using cumulative throughput with a similar experiments found in research literature.  In this experiment a given set of nodes communicate directly with each other using TCP connections.  We used same message type (46 bytes in serialized format) used for earlier experiment. The given cumulative throughput is at the application level message rate omitting the TCP overhead and all communications are through the network. For all experiments we use 20 message buffer.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{allthroughput.png}
        \caption{All to All Throughput}
        \label{allthroughput}
\end{figure}

As shown in Figure \ref{allthroughput} system linearly scales with the number of nodes.  Compared to Naiad and Storm we have achieved better performance.  Figure \ref{alllatency} shows the mean message latency for all messages. We measured the message message latency during the same experiment by sending the time stamp along with each message. Compared to Storm we have achieved much lesser latency level. Naiad reports very low latency. However they do not measure the latency of the data messages in the stressed system but measure the latency of control messages.  MillWheel has reported very low latency level compared to our results. Further it is worth note that the latency can be reduced by reducing the buffer size and hence reducing the throughput. We choose 20 message buffer since it gives better throughput as well as comparable low latency.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{alllatency.png}
        \caption{All to All Latency}
        \label{alllatency}
\end{figure}

\subsection{Fault tolerance}
For this experiment we analyzed the resilience of the system in the presence of node failures. Initially we had a load balanced system where the event producer sends events to 8 receiving nodes. The producer produced messages at the rate of around 2500000 and each receiving node receive messages at a rate around 320000 ($\sim$ 2500000 / 8). Then we shutdown one process by one process within 1 minute interval. This increased the throughput of each receiving node. Figure \ref{rthroughput} shows the throughput received at the last node to shutdown with the time. As depicts in the Figure \ref{rthroughput} when a node fails system automatically removes the node from its configurations and route the messages to other nodes balancing load among existing ones.

\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{failure_throughput.png}
        \caption{Receiver Throughput}
        \label{rthroughput}
\end{figure}

### 2015-02-10 paper/system_architecture.tex
rewritting those chapters

\section{System Design}
In this section we elaborate on our underlying system that is capable of processing 2.4 million messages per second. We emphasize three features of this design. Firstly between any two nodes there is a set of TCP connections to send messages. All processes share the same set of connections, which we cumulatively call the bridge. The advantage of this approach compared to having TCP connections per process is that it utilizes all available bandwidth even there is only one process per node. However since this is a common bridge now we have to send the receiving process id to dispatch the message, sending  process id to parse the message and sequence number to order messages. As we discuss later our experiments show that this approach perform several times better than the other systems despite the above overhead. Secondly we have implemented \textit{OutputStream} and \textit{InputStream} interfaces on top of java non blocking I/O API. Using these input and output streams we provide \textit{DataOutput} and \textit{DataInput} APIs to directly 
serialize message data to binary format without sending any metadata along the message. Finally except messages themselves, all the other objects used to serialize and parse the messages and parse the messages are not created at message processing time. All these objects are created when a process sends the first message and it is a one-time cost. Rest of this section further explains the design techniques we used to realize these concepts.
\subsection{Inter Process Communication}
Figure \ref{interprocess} shows the design involved in sending a message from one process to another. Lets assume there is a message at a client side process and that process wants to send this message to another process. First the client process sends this message to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to send and it push the message to all streams. Stream decides which node to send this message and passes the message along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} sends message to correct \textit{ClientConnection} using target node. \textit{ClientConnection} picks an available \textit{DataOutput} from its pool of \textit{DataOutputs} and invokes the serialize method of the message to send that to server side.

At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time). Once a \textit{ServerTask} receives a binary message, it creates the event using the sending process id to identify the event type. Then it passes this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id. Message ordering happens at the process level if required. This illustrates how we have shared a pool of TCP connections among processes and next we show how we have implemented \textit{OutputStream} and \textit{InputStream} with non blocking I/O API without  creating objects dynamically.
\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
\subsection{Client Side}
Figure \ref{client} shows how we have implemented \textit{OutputStream} interface using our \textit{DataWriter}. \textit{DataWriter} contains a \textit{ByteBuffer}. When a higher layer method writes data to \textit{DataOuput}, \textit{DataWriter} puts those bytes to its byte buffer. When a selection event occurs, \textit{Datawriter} writes its byte buffer to underline \textit{SocketChannel}. This way only one byte buffer is used as an intermediary to create an output stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{client.png}
        \caption{OutputStream Implementation}
        \label{client}
\end{figure}
\subsection{Server Side}
Figure \ref{server} shows how we have implemented \textit{InputStream} interface using our \textit{DataReader}. Similar to client side, \textit{DataReader} has a \textit{ByteBuffer}. When a selection event occurs \textit{DataReader} inputs bytes from the \textit{SocketChannel} to its byte buffer. When a \textit{ServerTask} reads data through \textit{DataInput}, it reads data from the byte buffer and passes to higher layer. In this way we have implemented an input stream with non blocking I/O API.
\begin{figure}[!t]
        \centering
        \includegraphics[width=3.0in]{server.png}
        \caption{InputStream Implementation}
        \label{server}
\end{figure}

>>>>

\section{System Design}
In this section we describe key aspects of our system design which increase the throughput of inter node communication. First of all it is worth to note that even with one TCP connection, it is possible to saturate the network link usage if the sending and receiving processes send and receive byte arrays without a delay. However serializing java object messages to byte arrays and deserializing back takes considerable amount of time. Therefore we can improve the communication performance by reducing this message processing time. We use multitasking and bulk message serializing to solve this problem.

In this design, we use thread pools both at client side and server side to processes messages while sharing a single TCP connection. Firstly both thread pools access underline connection resource only to send receive byte arrays making message serialization and deserialization processes parallel. Secondly it serializes messages together to improve serialization process.  Figure \ref{interprocess} shows the design involved in sending a message from one process to another.

First the client process thread sends this message to its \textit{ElementContainer}. \textit{ElementContainer} holds all the streams (this is an abstraction of a link in the process graph) to which this message needs to be send and it pushes the message to all the streams. Stream decides the target node and buffers the messages for a configured amount of messages. Then it passes the buffered  message list along the target node details to \textit{ConnectionManager} which holds \textit{ClientConnection}s for each node. \textit{ConnectionManager} serializes the message list and sends the byte array to correct \textit{ClientConnection} using the target node. \textit{ClientConnection} uses the \textit{DataOutput} which wraps the TCP connection to send message to server side.

At the server side there is a set of \textit{ServerTask}s which reads receiving messages from a pool of \textit{DataInput} streams available in the \textit{ServerConnection} (we register these connections at the connections creating time) and send them to receiving processes. First a server task acquires a \textit{DataInput} stream and reads the binary message and release it. Then it deserializes the binary message using the sending process id to identify the event type.  After that it passes this message to the \textit{WorkerContainer} which holds all processes. Finally \textit{WorkerContainer} dispatches this message to correct process using receiving process id. Message ordering happens at the process level if required.

\begin{figure}[!tii]
        \centering
        \includegraphics[width=3.0in]{interprocess.png}
        \caption{Communication between two Process}
        \label{interprocess}
\end{figure}
